---
title: "First Exploration of PILA Data"
author: "Lennard Naumann"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(tidyverse)
library(stringi)
library(ggExtra)

#### Data Prep Jobs

data <- read.csv("../D4 Third Download/pila_2023-04-24_2011_2022_1_vars.csv", encoding = "latin1") %>% 
  mutate(
    Departamento = stri_trans_general(Departamento, id = "Latin-ASCII"),
    sector = stri_trans_general(sector, id = "Latin-ASCII")
  ) %>% 
  # Start in 2011, barely anything before that (see graphs w/o this restriction for comparison)
  filter(year >= 2011, sector == "B - EXPLOTACION DE MINAS Y CANTERAS") %>% 
  group_by(Departamento, year) %>% 
  summarize(value = sum(value, na.rm = T)) %>% 
  mutate(Departamento = str_to_upper(str_remove_all(Departamento, "[0-9]* - "))) %>% 
  filter(Departamento != "-NO DEFINIDO" & Departamento != "ARCHIPIELAGO DE SAN ANDRES, PROVIDENCIA Y SANTA CATALINA")

data_nz <- data %>% 
  group_by(Departamento) %>% 
  filter(sum(!is.na(value)) > 0) %>% 
  ungroup %>% 
  mutate(value = ifelse(is.na(value), 0, value))

data_sum <- data %>% 
  group_by(Departamento) %>% 
  summarise(
    n = sum(!is.na(value)),
    mean = mean(value, na.rm = T),
    sd = sd(value, na.rm = T)
  )


data_nz2 <- filter(data_nz, Departamento != "11 - Bogota, D.C.")

data_sum_nte <- data_sum
Munips_top75 <- data_sum_nte %>% 
  filter(mean >= quantile(data_sum_nte$mean, .25, na.rm = T)) %>% 
  pull(Departamento)
Munips_top05 <- data_sum_nte %>% 
  filter(mean >= quantile(data_sum_nte$mean, .95, na.rm = T)) %>% 
  arrange(-mean) %>% 
  pull(Departamento)
data_nte <- data_nz %>% 
  filter(Departamento %in% Munips_top75)

Munips_top75_full <- data_sum_nte %>% 
  filter(n == 12) %>% 
  pull(Departamento)
data_nte_full <- filter(data_nte, Departamento %in% Munips_top75_full)

#### Data prep Mining

anm <- read.csv("../../0Data/ANM/Mineria_-_Materiales_extraidos_en_Colombia_2012_a_2017.csv")
names(anm) <- stri_trans_general(names(anm), id = "Latin-ASCII")

dt_min <- anm %>% 
  filter(Recurso.Natural == "CARBON") %>% 
  transmute(
    Departamento = str_to_upper(stri_trans_general(Departamento, id = "Latin-ASCII")),
    year = as.numeric(Ano.Produccion),
    CoalOutput = ifelse(
      Unidad.Medida == "TONELADAS", Cantidad.Produccion, ifelse(
        Unidad.Medida == "LIBRAS", Cantidad.Produccion/2000, ifelse(
          Unidad.Medida == "GRAMOS", Cantidad.Produccion/1000000, ifelse(
            Unidad.Medida == "QUILATES", Cantidad.Produccion/5000000, NA
          ))))
  ) %>% 
  filter(!is.na(CoalOutput) & Departamento!="" & !is.na(Departamento)) %>%
  group_by(Departamento, year) %>% 
  dplyr::summarise(CoalOutput = sum(CoalOutput, na.rm = T)) %>% 
  ungroup()

# Create merged data
merged_data <- data_nte %>%
  transmute(
    Departamento = Departamento,
    year = year,
    NumeroTotalCotizantes = value
  ) %>% 
  left_join(dt_min, by = c("Departamento", "year"))
```



## Mining Jobs

### Path plot for total number of cotizantes

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte, aes(x = year, y = value, group = Departamento)) +
  geom_path(col = "lightgrey") +
  theme_minimal()

```


### Path plot for total number of cotizantes, only full observations

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte_full, aes(x = year, y = value, group = Departamento)) +
  geom_path(col = "lightgrey") +
  theme_minimal()

```


### Path plot for log(total number of cotizantes), only full observations

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte_full, aes(x = year, y = value, group = Departamento)) +
  geom_path(col = "lightgrey")+
  theme_minimal() +
  scale_y_log10()

```


### Total number of cotizantes by department

```{r, fig.height=12, fig.width=12, echo = F, results = F}
data_sum_nte %>% 
  mutate(Departamento = factor(
    x = Departamento,
    levels = (data_sum_nte %>% arrange(mean) %>% pull(Departamento))
  )) %>% 
  

  ggplot(., aes(y = Departamento, x = mean)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal()

```


### Distribution of log(total number of cotizantes)

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_sum_nte, aes(x = mean)) +
  geom_histogram() +
  theme_minimal() +
  scale_x_log10() +
  labs(x = "Mean no. of cotizantes per year")

```

### Distribution observed years per departamento

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_sum_nte, aes(x = n)) +
  geom_histogram(bins = 12) +
  theme_minimal() +
  labs(x = "Number of non-NA years")

```

### Scatter plot of coal output (ANM) and total number of cotizantes (PILA), logged axes

```{r, fig.height=12, fig.width=12, echo = F, results = F}
graph <- ggplot(merged_data) +
    geom_point(size = 5, aes(x = CoalOutput, y = NumeroTotalCotizantes, col = year, shape = Departamento)) +
    geom_smooth(col = "black", aes(x = CoalOutput, y = NumeroTotalCotizantes)) +
    scale_y_log10() + scale_x_log10() +
    scale_shape_manual(values = 1:27) +
theme(legend.position = "bottom")

ggMarginal(graph, type = "boxplot")
```

### Scatter plot of coal output (ANM) and total number of employees (PILA), without top and bottom 5%-quantiles, logged axes

```{r, fig.height=12, fig.width=12, echo = F, results = F}
graph2 <- merged_data %>% 
  filter(
    NumeroTotalCotizantes > quantile(merged_data$NumeroTotalCotizantes, .05, na.rm = T),
    NumeroTotalCotizantes < quantile(merged_data$NumeroTotalCotizantes, .95, na.rm = T),
    CoalOutput > quantile(merged_data$CoalOutput, .05, na.rm = T),
    CoalOutput < quantile(merged_data$CoalOutput, .95, na.rm = T)
  ) %>% 
  
  ggplot(.) +
  geom_point(size = 5, aes(x = CoalOutput, y = NumeroTotalCotizantes, col = year, shape = Departamento)) +
  geom_smooth(col = "black", aes(x = CoalOutput, y = NumeroTotalCotizantes)) +
  scale_y_log10() + scale_x_log10() +
  scale_shape_manual(values = 1:27) +
  theme(legend.position = "bottom")

ggMarginal(graph2, type = "boxplot")

```
