---
title: "Firs Exploration of PILA Data"
author: "Lennard Naumann"
date: "2023-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(tidyverse)
library(stringi)
library(ggExtra)

#### Data Prep Jobs

data <- read.csv("../D4 Third Download/pila_2023-04-24_2011_2022_1_vars.csv", encoding = "latin1") %>% 
  mutate(
    Municipio = stri_trans_general(Municipio, id = "Latin-ASCII"),
    sector = stri_trans_general(sector, id = "Latin-ASCII")
  ) %>% 
  # Start in 2011, barely anything before that (see graphs w/o this restriction for comparison)
  filter(year >= 2011, sector == "B - EXPLOTACION DE MINAS Y CANTERAS")

data_nz <- data %>% 
  group_by(Municipio) %>% 
  filter(sum(!is.na(value)) > 0) %>% 
  ungroup %>% 
  mutate(value = ifelse(is.na(value), 0, value))

data_sum <- data %>% 
  group_by(Municipio) %>% 
  summarise(
    n = sum(!is.na(value)),
    mean = mean(value, na.rm = T),
    sd = sd(value, na.rm = T)
  )


data_nz2 <- filter(data_nz, Municipio != "11001 - Bogot, D.C.", Municipio != "76001 - Cali")

data_sum_nte <- data_sum
Munips_top75 <- data_sum_nte %>% 
  filter(mean >= quantile(data_sum_nte$mean, .25, na.rm = T)) %>% 
  pull(Municipio)
Munips_top05 <- data_sum_nte %>% 
  filter(mean >= quantile(data_sum_nte$mean, .95, na.rm = T)) %>% 
  arrange(-mean) %>% 
  pull(Municipio)
data_nte <- data_nz %>% 
  filter(Municipio %in% Munips_top75)

Munips_top75_full <- data_sum_nte %>% 
  filter(n == 8) %>% 
  pull(Municipio)
data_nte_full <- filter(data_nte, Municipio %in% Munips_top75_full)

#### Data prep Mining

anm <- read.csv("../../0Data/Colombia ANM/Mineria_-_Materiales_extraidos_en_Colombia_2012_a_2017.csv")
names(anm) <- stri_trans_general(names(anm), id = "Latin-ASCII")

dt_min <- anm %>% 
  filter(Recurso.Natural == "CARBON") %>% 
  transmute(
    Municipio = str_to_upper(stri_trans_general(Municipio.Productor, id = "Latin-ASCII")),
    Departamento = str_to_upper(stri_trans_general(Departamento, id = "Latin-ASCII")),
    year = as.numeric(Ano.Produccion),
    CoalOutput = ifelse(
      Unidad.Medida == "TONELADAS", Cantidad.Produccion, ifelse(
        Unidad.Medida == "LIBRAS", Cantidad.Produccion/2000, ifelse(
          Unidad.Medida == "GRAMOS", Cantidad.Produccion/1000000, ifelse(
            Unidad.Medida == "QUILATES", Cantidad.Produccion/5000000, NA
          ))))
  ) %>% 
  filter(!is.na(CoalOutput) & Municipio!="") %>% 
  
  # Correct non-matching municipality names:
  mutate(
    # Corrections by hand
    Municipio = case_when(
      Municipio == "AMAGA" ~ "AMAG",
      Municipio == "ANGELOPOLIS" ~ "ANGELPOLIS",
      Municipio == "AGUSTIN CODAZZI" ~ "NA",
      Municipio == "ALBANIA - SANTANDER" ~ "NA",
      Municipio == "ALBANIA - LA GUAJIRA" ~ "NA",
      Municipio == "AMAGA" ~ "AMAG",
      Municipio == "ARBOLEDAS - NORTE DE SANTANDER" ~ "ARBOLEDAS",
      Municipio == "BETEITIVA" ~ "BETITIVA",
      Municipio == "BOYACA - BOYACA" ~ "BOYAC",
      Municipio == "BRICENO - BOYACA" ~ "BRICEO",
      Municipio == "CACOTA" ~ "CCOTA",
      Municipio == "CAJIBIO" ~ "CAJIC",
      Municipio == "CAPARRAPI" ~ "CAPARRAP",
      Municipio == "CHINACOTA" ~ "CHINCOTA",
      Municipio == "CHIQUINQUIRA" ~ "CHIQUINQUIR",
      Municipio == "CHIRIGUANA" ~ "CHIRIGUAN",
      Municipio == "CHISCAS" ~ "NA",
      Municipio == "CHITAGA" ~ "CHITAG",
      Municipio == "CHIVATA" ~ "CHIVAT",
      Municipio == "CIENAGA - MAGALENA" ~ "NA",
      Municipio == "COMBITA" ~ "CMBITA",
      Municipio == "CUCUNUBA" ~ "CUCUNUB",
      Municipio == "CUCUTA" ~ "CUCAITA",
      Municipio == "EL CARMEN DE CHUCURI" ~ "EL CARMEN DE CHUCUR",
      Municipio == "EL ESPINO" ~ "NA",
      Municipio == "EL TAMBO - CAUCA" ~ "EL TAMBO",
      Municipio == "GUACHETA" ~ "GUACHET",
      Municipio == "GUAVATA" ~ "GUATAVITA",
      Municipio == "HERRAN" ~ "HERRN",
      Municipio == "JAMUNDI" ~ "JAMUND",
      Municipio == "JERICO - BOYACA" ~ "JERIC",
      Municipio == "JERUSALEN" ~ "NA",
      Municipio == "LANDAZURI" ~ "LANDZURI",
      Municipio == "MONGUI" ~ "MONGU",
      Municipio == "MORALES - CAUCA" ~ "MORALES",
      Municipio == "MUTISCUA" ~ "NA",
      Municipio == "NEMOCON" ~ "NEMOCN",
      Municipio == "NUEVO COLON" ~ "NA",
      Municipio == "PACHAVITA" ~ "NA",
      Municipio == "PATIA" ~ "PATA",
      Municipio == "PAZ DE RIO" ~ "PAZ DE RO",
      Municipio == "PESCA" ~ "NA",
      Municipio == "RAQUIRA" ~ "NA",
      Municipio == "SABOYA" ~ "SABOY",
      Municipio == "SAMACA" ~ "SAMAC",
      Municipio == "SAN CAYETANO - CUNDINAMARCA" ~ "SAN CAYETANO",
      Municipio == "SAN CAYETANO - NORTE DE SANTANDER" ~ "SAN CAYETANO",
      Municipio == "SAN JOSE" ~ "NA",
      Municipio == "SAN JOSE DE MIRANDA" ~ "NA",
      Municipio == "SAN MARTIN - META" ~ "SAN MARTN",
      Municipio == "SAN MIGUEL - SANTANDER" ~ "SAN MIGUEL",
      Municipio == "SANTA MARIA - BOYACA" ~ "NA",
      Municipio == "SANTIAGO - NORTE DE SANTANDER" ~ "SANTIAGO",
      Municipio == "SANTO TOMAS" ~ "NA",
      Municipio == "SOCOTA" ~ "SOCOT",
      Municipio == "SOPETRAN" ~ "SOPETRN",
      Municipio == "SOTAQUIRA" ~ "SOTAQUIR",
      Municipio == "SUAREZ - CAUCA" ~ "SUREZ",
      Municipio == "SUSACON" ~ "SUSA",
      Municipio == "TIBANA" ~ "NA",
      Municipio == "TITIRIBI" ~ "TITIRIB",
      Municipio == "TIBU" ~ "TIB",
      Municipio == "TOLEDO - ANTIOQUIA" ~ "TOLEDO",
      Municipio == "TOLEDO - NORTE DE SANTANDER" ~ "TOLEDO",
      Municipio == "TOPAGA" ~ "TPAGA",
      Municipio == "TURMEQUE" ~ "TURMEQU",
      Municipio == "VELEZ" ~ "NA",
      Municipio == "VENECIA - ANTIOQUIA" ~ "VENECIA",
      Municipio == "VILLAPINZON" ~ "VILLAPINZN",
      Municipio == "YACOPI" ~ "YACOP",
      Municipio == "ZIPAQUIRA" ~ "ZIPAQUIR",
      T ~ Municipio
    )
  ) %>% 
  filter(Municipio != "NA") %>% 
  group_by(Departamento, Municipio, year) %>% 
  dplyr::summarise(CoalOutput = sum(CoalOutput, na.rm = T)) %>% 
  ungroup() %>% 
  select(-Departamento)

# Create merged data
merged_data <- data_nte %>%
  transmute(
    Municipio = str_to_upper(str_remove_all(Municipio, "[0-9]* - ")),
    year = year,
    NumeroTotalEmpleados = value
  ) %>% 
  left_join(dt_min, by = c("Municipio", "year"))
```



## Mining Jobs

### Path plot for total number of cotizantes

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte, aes(x = year, y = value, group = Municipio)) +
  geom_path(col = "lightgrey") +
  theme_minimal()

```


### Path plot for total number of cotizantes, only full observations

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte_full, aes(x = year, y = value, group = Municipio)) +
  facet_wrap(variable~.)+
  geom_path(col = "lightgrey") +
  theme_minimal()

```


### Path plot for log(total number of cotizantes), only full observations

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_nte_full, aes(x = year, y = value, group = Municipio)) +
  geom_path(col = "lightgrey")+
  theme_minimal() +
  scale_y_log10()

```


### Total number of cotizantes by municipality

```{r, fig.height=12, fig.width=12, echo = F, results = F}
data_sum_nte %>% 
  mutate(Municipio = factor(
    x = ifelse(!(Municipio %in% Munips_top05), "Other", Municipio),
    levels = c(Munips_top05, "Other")
  )) %>% 
  group_by(Municipio) %>% 
  summarise(mean = sum(mean, na.rm = T), placeholder = "") %>% 
  ungroup() %>% 

  ggplot(., aes(y = Municipio, x = mean)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal()

```


### Distribution of log(total number of cotizantes)

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_sum_nte, aes(x = log(mean))) +
  geom_histogram() +
  theme_minimal() +
  labs(x = "Log(mean no. of cotizantes per year)")

```

### Distribution observed years per municipality

```{r, fig.height=12, fig.width=12, echo = F, results = F}
ggplot(data_sum_nte, aes(x = n)) +
  geom_histogram() +
  theme_minimal() +
  labs(x = "Number of non-NA years")

```

### Scatter plot of coal output (ANM) and total number of cotizantes (PILA)

```{r, fig.height=12, fig.width=12, echo = F, results = F}
graph <- ggplot(merged_data, aes(x = CoalOutput, y = NumeroTotalEmpleados, col = year)) +
  geom_point() +
  geom_smooth(col = "black") +
  scale_y_log10() + scale_x_log10()

ggMarginal(graph, type = "boxplot")
```

### Scatter plot of coal output (ANM) and total number of employees (PILA), without top and bottom 5%-quantiles

```{r, fig.height=12, fig.width=12, echo = F, results = F}
graph2 <- merged_data %>% 
  filter(
    NumeroTotalEmpleados > quantile(merged_data$NumeroTotalEmpleados, .05, na.rm = T),
    NumeroTotalEmpleados < quantile(merged_data$NumeroTotalEmpleados, .95, na.rm = T),
    CoalOutput > quantile(merged_data$CoalOutput, .05, na.rm = T),
    CoalOutput < quantile(merged_data$CoalOutput, .95, na.rm = T)
  ) %>% 
  
  ggplot(., aes(x = CoalOutput, y = NumeroTotalEmpleados, col = year)) +
  geom_point() +
  geom_smooth(col = "black") +
  scale_y_log10() + scale_x_log10()

ggMarginal(graph2, type = "boxplot")

```
